When Subgraph Isomorphism is Really Hard, and Why This Matters for Graph Databases

### 2018-02-14 paper/paper.tex
Final sentence
Explain T_io (review 1)

This equation is repeated and expanded upon by \citet{DBLP:journals/tods/YanYH05} to explicitly
separate I/O and isomorphism testing costs, obtaining a query response time of \[
    T_{\mathit{search}} + \left|C_q\right| \cdot (T_{\mathit{io}} + T_{\mathit{iso\_test}})
    \textnormal{.} \] The authors explicitly state that ``the value of $T_{\mathit{iso\_test}}$ does
not change much for a given query''. The argument presented is that
\begin{displayquote}``Sequential scan is very costly
because one has to not only access the whole graph database but also check subgraph isomorphism. It
is known that subgraph isomorphism is an \NP-complete problem. Clearly, it is necessary to build
graph indices in order to help processing graph queries.''\end{displayquote}
We reassess these claims in light of
what we now know about the behaviour of modern subgraph isomorphism algorithms and the nature
of solving \NP-complete problems. Contrary to the authors' claims,
we do not expect $T_{\mathit{iso\_test}}$ to be anything like a constant, even if the orders of the
input graphs are similar. In particular, any instance which can be excluded based upon filtering
must have a very small proof of unsatisfiability. These instances \emph{should} be trivial with any
reasonably good subgraph isomorphism algorithm. Thus, all filtering \emph{should} be doing is eliminating the startup
costs of a trivial subgraph isomorphism call. The fact that filtering was successful empirically
should make us wonder whether the subgraph isomorphism algorithm used was substantially
weaker than the state of the art.
Indeed, the algorithm is described only as ``the simplest approach'' in the
paper.  Additionally, the experiments focus on reducing the size of the candidate set, without
considering the time taken to verify different candidate set instances.  The claim that
$T_{\mathit{iso\_test}}$ does not change much is not justified experimentally, and no consideration
is given as to whether this would hold true for other subgraph isomorphism algorithms.

@@

Of course, hard random instances of the kind we describe in
\cref{section:non-induced,section:induced} are also atypical, and should not form the sole basis for
benchmarking. In particular, distance-based filtering
\citep{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16} has little
to no effect on these instances, but is extremely beneficial on other kinds of graph, including
those from real-world applications.  These graphs also have a low degree spread and very little
structure to exploit, which could mislead experimenters as to the value of certain kinds of
inference. \citet{o:GentW95} give an example of an exam timetabling graph which contains an
unexpected ten-vertex clique, which would be extremely unlikely in a randomly generated graph with a
similar order and density.

>>>>

This equation is repeated and expanded upon by \citet{DBLP:journals/tods/YanYH05} to explicitly
separate I/O and isomorphism testing costs into $T_{\mathit{io}}$ and $T_{\mathit{iso\_test}}$
respectively, obtaining a query response time of \[
    T_{\mathit{search}} + \left|C_q\right| \cdot (T_{\mathit{io}} + T_{\mathit{iso\_test}})
    \textnormal{.} \] The authors explicitly state that ``the value of $T_{\mathit{iso\_test}}$ does
not change much for a given query''. The argument presented is that
\begin{displayquote}``Sequential scan is very costly
because one has to not only access the whole graph database but also check subgraph isomorphism. It
is known that subgraph isomorphism is an \NP-complete problem. Clearly, it is necessary to build
graph indices in order to help processing graph queries.''\end{displayquote}
We reassess these claims in light of
what we now know about the behaviour of modern subgraph isomorphism algorithms and the nature
of solving \NP-complete problems. Contrary to the authors' claims,
we do not expect $T_{\mathit{iso\_test}}$ to be anything like a constant, even if the orders of the
input graphs are similar. In particular, any instance which can be excluded based upon filtering
must have a very small proof of unsatisfiability. These instances \emph{should} be trivial with any
reasonably good subgraph isomorphism algorithm. Thus, all filtering \emph{should} be doing is eliminating the startup
costs of a trivial subgraph isomorphism call. The fact that filtering was successful empirically
should make us wonder whether the subgraph isomorphism algorithm used was substantially
weaker than the state of the art.
Indeed, the algorithm is described only as ``the simplest approach'' in the
paper.  Additionally, the experiments focus on reducing the size of the candidate set, without
considering the time taken to verify different candidate set instances.  The claim that
$T_{\mathit{iso\_test}}$ does not change much is not justified experimentally, and no consideration
is given as to whether this would hold true for other subgraph isomorphism algorithms.

@@

Of course, hard random instances of the kind we describe in
\cref{section:non-induced,section:induced} are also atypical, and should not form the sole basis for
benchmarking. In particular, distance-based filtering
\citep{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16} has little
to no effect on these instances, but is extremely beneficial on other kinds of graph, including
those from real-world applications.  These graphs also have a low degree spread and very little
structure to exploit, which could mislead experimenters as to the value of certain kinds of
inference. \citet{o:GentW95} give an example of an exam timetabling graph which contains an
unexpected ten-vertex clique, which would be extremely unlikely in a randomly generated graph with a
similar order and density. We therefore emphasise the importance of working with a diverse set of
benchmark instances, and of performing a careful family-by-family analysis of the results.
